Measurement of nuclear excitation functions for proton induced reactions (E$_{\text{p}}$ = 40 -- 90 MeV) on natural Nb and Cu

### 2018-02-03 nb_px_paper/nb_px_paper.tex
Started merging Stephen's 1st round revisions.

A stacked target of natural Nb, Cu, and Al thin foils were irradiated by 100 MeV protons at the Los Alamos National Laboratory's Isotope Production Facility.
The irradiation was motivated by interest in the characterization of the \ce{^{93}Nb}(p,4n)\ce{^{90}Mo} reaction as a new dosimetry standard for intermediate-energy protons.
Through this, a set of 38 measurements of cross sections for the \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions between 40.34 - 89.74 MeV are presented in this work, as well as 5 independent measurements of isomer branching ratios.
The cross sections were measured using the well-characterized \ce{^{nat}Al}(p,x)\ce{^{22}Na}, \ce{^{nat}Al}(p,x)\ce{^{24}Na}, \ce{^{nat}Cu}(p,x)\ce{^{56}Co}, \ce{^{nat}Cu}(p,x)\ce{^{62}Zn}, and \ce{^{nat}Cu}(p,x)\ce{^{65}Zn} monitor reactions  for accurate proton fluence measurement, with all activities quantified using high-purity germanium spectroscopy.
The measured cross sections are shown to be in excellent agreement with  literature values, and have been measured with significantly less uncertainty than previous measurements, through the use of variance minimization techniques to reduce proton energy uncertainties.
This work reports the first measurements of the  \ce{^{nat}Nb}(p,x)\ce{^{82m}Rb} reaction, as well as the first measurement of the independent cross sections for    \ce{^{nat}Cu}(p,x)\ce{^{52\text{m}}Mn}, \ce{^{nat}Cu}(p,x)\ce{^{52\text{g}}Mn}, and \ce{^{nat}Nb}(p,x)\ce{^{85\text{g}}Y} in the 40 - 90 MeV region.
Comparisons against the reaction modeling codes EMPIRE and TALYS are used to explore the impact of pre-equilibrium particle emission in the reaction dynamics for 40 - 90 MeV \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions.

@@

Every year, approximately 17 million nuclear medicine procedures (both diagnostic and therapeutic) are performed in the U.S. alone - a multi-billion dollar industry which has made incredible strides in improving our ability to detect and treat a variety of life-threatening diseases \cite{Delbeke2011,NSACIsotopesSubcommittee2015}.
The vast majority of the radioisotopes currently used for these procedures are produced in the field's array of low- (E \textless\ 30 MeV / A) and intermediate-energy (30 \textless\ E \textless\ 200 MeV / A) accelerator capabilities, which routinely produce many of the staple medical radionuclides, such as \ce{^{18}F} \ce{^{68}Ge}, \ce{^{82}Rb}, and \ce{^{123}I}, as well as many of the non-medical radioisotopes of commercial value, such as  \ce{^{32}Si}, \ce{^{73}As}, \ce{^{95m}Tc}, and  \ce{^{109}Cd} \cite{international2009iaea,schlyer2008cyclotron}. 
\comment{Should \enquote{intermediate-energy} be less than 200 MeV, or 100 MeV?  I find conflicting descriptions in the literature.\\Lee:  I believe that 60 MeV is considered as the transition point\\ Stephen: }
 The future of nuclear medicine would appear to be the paradigm of personalized medicine - targeted radionuclide therapy to spare healthy tissue \cite{Mulford2005,Qaim201731}, and theranostic medicine, which pairs an imaging isotope with a therapeutic isotope of the same element, to provide simultaneous, real-time dose delivery and verification, leading to drastic reductions in prescribed patient dose \cite{Muller2014,Bentzen2005,Srivastava2012}. 
Candidate isotopes to fill these needs have been identified based on their decay properties \cite{NSACIsotopesSubcommittee2015,Qaim201731,bernstein2015nuclear}, and a series of campaigns are underway to perform targeted, high-priority measurements of thin-target cross sections and thick-target integral yields, to facilitate production in sufficient quantities for cell studies.

One significant obstacle to both high-fidelity measurements of production cross sections for emerging medical radionuclides, as well as conventional isotope production, is a lack of well-characterized  dosimetry standards.  
This is particularly true for higher-energy charged particle beams, where there is currently a paucity of such data. 
Indeed, the development of new dosimetry standards and the improved evaluation of existing standards is one of the areas of greatest cross-cutting need for nuclear data \cite{bernstein2015nuclear}. 
Charged particle dosimetry data currently exists for low-to-intermediate energy charged particle beams (E \textless\ 50 MeV / A), but experimental data used for this evaluation is sparse above approximately 30 MeV / A and  uncertainties in experimental cross sections are large (6-15\%) \cite{gul2001charged}. 
While work is needed to improve upon existing dosimetry data, the development of new reactions can expand the available range of options for the monitoring of charged particle beams.

>>>>

A stacked target of natural Nb, Cu, and Al thin foils were irradiated by 100 MeV protons at the Los Alamos National Laboratory's Isotope Production Facility.
The irradiation was motivated by interest in the characterization of the \ce{^{93}Nb}(p,4n)\ce{^{90}Mo} reaction as a new dosimetry standard for intermediate-energy protons.
A set of 38 measurements of cross sections for the \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions between 40.34 -- 89.74 MeV are presented in this work, as well as 5 independent measurements of isomer branching ratios.
Cross sections were measured using the well-characterized \ce{^{nat}Al}(p,x)\ce{^{22}Na}, \ce{^{nat}Al}(p,x)\ce{^{24}Na}, \ce{^{nat}Cu}(p,x)\ce{^{56}Co}, \ce{^{nat}Cu}(p,x)\ce{^{62}Zn}, and \ce{^{nat}Cu}(p,x)\ce{^{65}Zn} monitor reactions  for accurate proton fluence measurement, with all activities quantified using high-purity germanium spectroscopy.
Variance minimization techniques were employed to reduce systematic uncertainties in proton energy and fluence, improving the reliability of these measurements. 
The measured cross sections are shown to be in excellent agreement with  literature values, and have been measured with improved precision compared with previous measurements.
This work reports the first measurements of the  \ce{^{nat}Nb}(p,x)\ce{^{82m}Rb} reaction, as well as the first measurement of the independent cross sections for    \ce{^{nat}Cu}(p,x)\ce{^{52\text{m}}Mn}, \ce{^{nat}Cu}(p,x)\ce{^{52\text{g}}Mn}, and \ce{^{nat}Nb}(p,x)\ce{^{85\text{g}}Y} in the 40 - 90 MeV region.
Comparisons against the reaction modeling codes EMPIRE and TALYS are used to explore the impact of pre-equilibrium particle emission in the reaction dynamics for 40 - 90 MeV \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions.

\comment{Stephen:  Not strictly true if you've omitted Na22 and Co56 from your fluence determination.  Update monitors here following discussion.}

\comment{Stephen: Unless you've already discussed this with co-authors, I would have someone from LANL (i.e. Meiring) instead or in addition.  Add one following discussion.}

@@

Every year, approximately 17 million nuclear medicine procedures (both diagnostic and therapeutic) are performed in the U.S. alone,  which has made incredible strides in improving our ability to detect and treat a variety of life-threatening diseases \cite{Delbeke2011,NSACIsotopesSubcommittee2015}.
The vast majority of the radioisotopes currently used for these procedures are produced in the field's array of low- (E \textless\ 30 MeV / A) and intermediate-energy (30 \textless\ E \textless\ 200 MeV / A) accelerator capabilities, which routinely produce many of the staple medical radionuclides, such as \ce{^{11}C}, \ce{^{18}F}, \ce{^{68}Ge}, \ce{^{82}Rb}, and \ce{^{123}I}, as well as many of the non-medical radioisotopes of commercial value, such as  \ce{^{32}Si}, \ce{^{73}As}, \ce{^{95m}Tc}, and  \ce{^{109}Cd} \cite{international2009iaea,schlyer2008cyclotron}. 
\comment{Should \enquote{intermediate-energy} be less than 200 MeV, or 100 MeV?  I find conflicting descriptions in the literature.\\Lee:  I believe that 60 MeV is considered as the transition point\\ Stephen: In this context, I would make this definition based on what machines in the US are routinely used for medical isotope production - based on this, I agree with 200 MeV. }
 The future of nuclear medicine would appear to be the paradigm of personalized medicine --- targeted radionuclide therapy to spare healthy tissue \cite{Mulford2005,Qaim201731}, and theranostic medicine, which pairs an imaging isotope with a therapeutic isotope (frequently, of the same element), to provide simultaneous, real-time dose delivery and verification, leading to drastic reductions in prescribed patient dose \cite{Muller2014,Bentzen2005,Srivastava2012}. 
Other variants of theranostic medicine exist, including pre-imaging for treatment planning, or delivery of a single compound with different radioelements for imaging/therapy where the inter-element biodistribution has been validated.  
Candidate isotopes to meet these needs have been identified based on their chemical and radioactive decay properties \cite{NSACIsotopesSubcommittee2015,Qaim201731,bernstein2015nuclear}, and a series of campaigns are underway to perform targeted, high-priority measurements of thin-target cross sections and thick-target integral yields.
These studies will serve to facilitate the production of pre-clinical quantities of radioactivity.

One significant obstacle to both high-fidelity measurements of production cross sections for emerging medical radionuclides, as well as conventional isotope production, is a lack of well-characterized  dosimetry standards.  
This is particularly true for higher-energy charged particle beams, where there is currently a paucity of such data. 
Indeed, the development of new dosimetry standards and the improved evaluation of existing standards is one of the areas of greatest cross-cutting need for nuclear data \cite{bernstein2015nuclear}. 
Charged particle dosimetry data currently exists for low-to-intermediate energy charged particle beams (E \textless\ 50 MeV / A), but experimental data used for this evaluation is sparse above approximately 30 MeV / A and  uncertainties in experimental cross sections are large (6-15\%) \cite{gul2001charged}. 
While work is needed to improve upon existing dosimetry data, the development of new reactions can expand the available range of options for the monitoring of charged particle beams.
\comment{Stephen: I'm having a hard time following this paragraph. Some of your syntax overlaps significantly with nuclear medicine \& medical physics terminology with different intended meanings. For example 'dosimetry' is typically taken to mean the measurement of absorbed energy per unit mass (J/kg) and associated measurements/specifications. Another example, 'dosimetry standard' is used quite often when referring to NIST air kerma reference standards which are the bedrock of Accredited Dosimetry Calibration Lab (ADCL) measurements. As you are more familiar with classic Nuclear Physics writing, perhaps you better know how this language will be interpreted. }
